Attrition in Online Learning: Understanding Persistence and Dropout in Massive Open Online Courses

### 2014-10-30 paper_v1.Rnw
More satisfaction results

<<loadcsv, echo=FALSE>>=
#library(RCurl)
#data <- getURL("https://raw.githubusercontent.com/whynotyet/MOOC_dropout/master/compiled_data_all_courses.csv", ssl.verifypeer=0L, followlocation=1L)
#a = read.csv(text=data)
a = read.csv("C:/Research/Analytics/pycode/user_code/halawa/publications/LAS2015/data_masters/compiled_data_all_courses.csv")

@@

First, we compared the distributions of satisfaction scores for learners who reported they stopped participating and those who reported not to have stopped. As figure \ref{fig:study1_satis_si_score_dist_comparison} shows, the mean satisfaction score ($mean=4.70$) for learners who reported to have stopped is 0.64 points lower than the other group ($mean=5.34$).

<<study1_satis_si_score_dist_comparison, echo=FALSE, fig.env='figure*', fig.pos='ht', fig.height=4, fig.width=9, fig.cap='Satisfaction index scores for learners who reported they stopped / did not stop participating before end of course'>>=
y0 = a$satisfaction[a$did_dropout == "No"]
y1 = a$satisfaction[a$did_dropout == "Yes"]
b = a[!is.na(a$did_dropout),]
ggplot(b) + geom_boxplot(aes(x=b$did_dropout,y=b$satisfaction)) + theme_bw() + labs(y="Satisfaction",x="Stopped participating?")
@

>>>>

<<loadcsv, echo=FALSE, warning=FALSE>>=
library(bitops)
library(RCurl)
data <- getURL("https://raw.githubusercontent.com/whynotyet/MOOC_dropout/master/compiled_data_all_courses.csv", ssl.verifypeer=0L, followlocation=1L)
a = read.csv(text=data)

@@

First, we compared the distributions of satisfaction scores for learners who reported they stopped participating and those who reported not to have stopped. A significant difference was observed ($t_{1469}=20.7, p \ll .0001, d=0.7$). The mean satisfaction score in the group that reported stopped (no stopped) participation was 4.70 (5.34).

<<study1_satis_si_score_dist_comparison, echo=FALSE, fig.env='figure*', fig.pos='ht', fig.height=4, fig.width=5, fig.cap='Satisfaction index scores for learners who reported they stopped / did not stop participating before end of course'>>=
#b = a[!is.na(a$did_dropout),]
#ggplot(b) + geom_boxplot(aes(x=b$did_dropout,y=b$satisfaction)) + theme_bw() + labs(y="Satisfaction",x="Stopped participating?")
@

We also examined differences in satisfaction between learners grouped by certain behaviors. First, we classified learners into 4 groups based on the fraction of course videos they viewed. The box plots for the satisfaction scores in each group are shown in figure \ref{fig:study1_satsfaction_by_vid_groups}. The mean satisfaction score increases slightly as the number of videos viewed by the learners increases, with the exception of the scores reported by learners in the 0-25\% group. Possible reasons for this are learners who join only to view a small part of the course or learners who join merely to explore.

<<study1_satsfaction_by_vid_groups, echo=FALSE, fig.env='figure*', fig.pos='ht', fig.height=4, fig.width=9, fig.cap='Satisfaction index scores for learners who reported they stopped / did not stop participating before end of course'>>=
b = a[!is.na(a$did_dropout),]
course_ids = unique(a$course_id)
course_num_videos = list()
for (cid in course_ids) {
  course_num_videos[[cid]] = quantile(a$num_videos_watched[a$course_id == cid], 0.95, na.rm=TRUE)
}
for(i in 1:nrow(b)) {
  x = b[i,]
  f = x[['num_videos_watched']]/course_num_videos[[x[['course_id']]]]
  if (is.na(f)) {level=NA}
  else if (f <= 0.25) {level = "0-25%"}
  else if  (f <= 0.50) {level = "25-50%"}
  else if  (f <= 0.75) {level = "50-75%"}
  else {level = "75-100%"}
  b[i,"fraction_videos_watched"] = level
}
b$fraction_videos_watched = factor(b$fraction_videos_watched)
c = b[!is.na(b$fraction_videos_watched),]
ggplot(c) + geom_boxplot(aes(x=c$fraction_videos_watched,y=c$satisfaction)) + theme_bw() + labs(y="Satisfaction",x="Fraction videos watched")
@
